---
title: "Prolific 2026-04-10"
date: last-modified
toc: true
format: 
  html:
    code-fold: true
    grid: 
      body-width: "800"
author: Jessica Helmer
knitr:
    opts_chunk: 
      dev: "ragg_png"
html-table-processing: none
---

The [](2026.01.14_prolific.qmd) document used data that I think was not recoded to align with scale direction. I renamed the values on Qualtrics, adding a `_r` tag to variables that *need* recoding. When I create recoded variables below, they will *not* have the `_r` tag. I also changed the BIS / BAS variable names to point specifically to their subscale (BIS or BAS).  

## Setup

::: panel-tabset
#### Setup
```{r}
#| message: false
#| code-fold: show
library(tidyverse)
library(patchwork)
library(gt)
library(rstan)

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = T)
```

#### Reading
```{r}
#| message: false
#| code-fold: show
dat_full <- qualtRics::read_survey(here::here("..", "Data", "prolific_2026.04.10.csv")) |>
  mutate(id = row_number(),
         .keep = "unused") |>
  filter(Progress == 100)
```

#### File Structure
```{r}
#| code-fold: show
if (!file.exists(here::here("Models"))) dir.create(here::here("Models"))
```
:::

## Cleaning

### Attention Checks

```{r}
#| column: margin
dat_full |>
  select(matches("attention")) |>
  imap(~ tibble(var = str_extract(.y, "([a-zA-Z-]+)_attention", group = 1),
                p = table(.x) |> prop.table() |> as_tibble()) |>
         unnest(cols = p)) |>
  list_rbind() |>
  rename(`Attention Check` = `.x`,
         p = n) |>
  gt(rowname_col = "var") |>
  fmt_percent(columns = p,
              decimals = 1) |>
  cols_label("p" = " ") |>
  tab_options(table.font.size = 12)
```


### Anxiety Scales

Cleaning, histograms, and single-factor CFA for each anxiety scale, with BIS and BAS separated.

::: {.callout-note collapse="true"}
#### Intolerance of Uncertainity

```{r}
IUS <- dat_full |>
  select(id, starts_with("IUS"), -matches("IUS_T"), -matches("IUS_attention")) |>
  mutate(across(!id, ~ recode_values(.x,
                                     "Entirely characteristic of me" ~ 5,
                                     "Very characteristic of me" ~ 4,
                                     "Somewhat characteristic of me" ~ 3,
                                     "A little characteristic of me" ~ 2,
                                     "Not at all characteristic of me" ~ 1)))

imap(IUS |> select(-id), \(data, var) tibble(data) |>
       ggplot(aes(x = data)) +
       geom_bar() +
       labs(x = var) +
       scale_x_continuous(limits = c(0.5, 5.5),
                          breaks = 1:5) + 
       scale_y_continuous(NULL,
                          breaks = NULL) +
       guides(x = guide_axis(cap = "both")) +
       theme_classic(base_size = 12) +
       theme(axis.line.y = element_blank())) |>
  wrap_plots()

IUS_model <- IUS |>
  select(-id) |>
  names() |>
  paste(collapse = " + ") |>
  paste("IUS =~ ", ... = _)

lavaan::cfa(model = IUS_model,
            data = IUS,
            estimator = "MLR") |>
  semPlot::semPaths(what = "std")
```
:::

::: {.callout-note collapse="true"}
#### Behavioral Inhibition Scale

```{r}
BIS <- dat_full |>
  select(id, starts_with("BIS"), -matches("BISBAS_T"), -matches("BISBAS_attention")) |>
  mutate(across(!id, ~ recode_values(.x,
                                     "Very true for me" ~ 4,
                                     "Somewhat true for me" ~ 3,
                                     "Somewhat false for me" ~ 2,
                                     "Very false for me" ~ 1))) |>
  mutate(across(ends_with("_r"), ~ 5 - .x,
                .names = "{str_replace(.col, '_r', '')}"),
         .keep = "unused")

imap(BIS |> select(-id), \(data, var) tibble(data) |>
       ggplot(aes(x = data)) +
       geom_bar() +
       labs(x = var) +
       scale_x_continuous(limits = c(0.5, 4.5),
                          breaks = 1:5) + 
       scale_y_continuous(NULL,
                          breaks = NULL) +
       guides(x = guide_axis(cap = "both")) +
       theme_classic(base_size = 12) +
       theme(axis.line.y = element_blank())) |>
  wrap_plots()

BIS_model <- BIS |>
  select(-id) |>
  names() |>
  paste(collapse = " + ") |>
  paste("BIS =~ ", ... = _)

lavaan::cfa(model = BIS_model,
            data = BIS,
            estimator = "MLR") |>
  semPlot::semPaths(what = "std")


# gr_BIS <- rstan::stan(file = here::here("Models", "graded-response_2pl_code.stan"),
#             data = list(K = 4,
#                         N_student = nrow(BIS),
#                         N_item = ncol(BIS) - 1, # taking out id column
#                         Y = BIS |> select(-id)))
# 
# gr_BIS |>
#   as.data.frame() |>
#   select(starts_with("kappa[1"))
# 
# gr_BIS |>
#   as.data.frame() |>
#   slice_sample(n = 50) |>
#   select(starts_with("a"), starts_with("b")) |>
#   mutate(rep = row_number()) |>
#   pivot_longer(!rep,
#                names_to = c("par", ".value"),
#                names_pattern = "([a-z]+)\\[(\\d)\\]") |>
#   pivot_longer(!c(rep, par),
#                names_to = "item", values_to = "est") |>
#   pivot_wider(names_from = par, values_from = est) |>
#   mutate(eta = )
```
:::

::: {.callout-note collapse="true"}
#### Behavioral Activation Scale

```{r}
BAS <- dat_full |>
  select(id, starts_with("BAS"), -matches("BISBAS_T"), -matches("BISBAS_attention")) |>
  mutate(across(!id, ~ recode_values(.x,
                                     "Very true for me" ~ 4,
                                     "Somewhat true for me" ~ 3,
                                     "Somewhat false for me" ~ 2,
                                     "Very false for me" ~ 1))) |>
  mutate(across(ends_with("_r"), ~ 5 - .x,
                .names = "{str_replace(.col, '_r', '')}"),
         .keep = "unused")

imap(BAS |> select(-id), \(data, var) tibble(data) |>
       ggplot(aes(x = data)) +
       geom_bar() +
       labs(x = var) +
       scale_x_continuous(limits = c(0.5, 4.5),
                          breaks = 1:5) + 
       scale_y_continuous(NULL,
                          breaks = NULL) +
       guides(x = guide_axis(cap = "both")) +
       theme_classic(base_size = 12) +
       theme(axis.line.y = element_blank())) |>
  wrap_plots()

BAS_model <- BAS |>
  select(-id) |>
  names() |>
  paste(collapse = " + ") |>
  paste("BAS =~ ", ... = _)

lavaan::cfa(model = BAS_model,
            data = BAS,
            estimator = "MLR") |>
  semPlot::semPaths(what = "std")
```
:::

::: {.callout-note collapse="true"}
#### State Trait Anxiety Inventory (Trait)

```{r}
STAI <- dat_full |>
  select(id, starts_with("STAI"), -matches("STAI-T_T"), -matches("STAI-T_attention")) |>
  mutate(across(!id, ~ recode_values(.x,
                                     "Almost never" ~ 4,
                                     "Often" ~ 3,
                                     "Sometimes" ~ 2,
                                     "Almost always" ~ 1))) |>
  mutate(across(ends_with("_r"), ~ 5 - .x,
                .names = "{str_replace(.col, '_r', '')}"),
         .keep = "unused") |>
  rename_with(~ str_replace(.x, "-T", ""),
              everything())

imap(STAI |> select(-id), \(data, var) tibble(data) |>
       ggplot(aes(x = data)) +
       geom_bar() +
       labs(x = var) +
       scale_x_continuous(limits = c(0.5, 4.5),
                          breaks = 1:5) + 
       scale_y_continuous(NULL,
                          breaks = NULL) +
       guides(x = guide_axis(cap = "both")) +
       theme_classic(base_size = 12) +
       theme(axis.line.y = element_blank())) |>
  wrap_plots()

STAI_model <- STAI |>
  select(-id) |>
  names() |>
  paste(collapse = " + ") |>
  paste("STAI =~ ", ... = _)

lavaan::cfa(model = STAI_model,
            data = STAI,
            estimator = "MLR") |>
  semPlot::semPaths(what = "std")
```
:::

### Decisions

```{r}
#| code-fold: show
dat <- dat_full |>
  select(id, matches("^[F][GL][PV]_[0-9\\.-]+$")) |>
  pivot_longer(!id,
               names_to = "item", values_to = "response") |>
  filter(!is.na(response)) |>
  separate_wider_regex(item, c(cond = "[F][GL][PV]", "_", rest = ".*")) |>
  separate_wider_regex(rest, c(fixed = "^\\d*", "-*", varied = ".*")) |>
  separate_wider_position(cond, c(1, "frame" = 1, "target" = 1)) |>
  mutate(.by = c(frame, target, fixed),
         item = cur_group_id(),
         .after = id) |>
  mutate(.by = c(item, id),
         trial = row_number(),
         .after = id) |>
  mutate(varied = as.numeric(varied),
         fixed = as.numeric(fixed),
         lottery = ifelse(target == "P", varied, fixed),
         certain = ifelse(target == "V", varied, fixed),
         ev_lottery = ifelse(frame == "L", 100 - lottery, lottery),
         ev_certain = ifelse(frame == "L", 100 - certain, certain),
         
         lottery_std = (lottery - mean(lottery)) / sd(lottery),
         certain_std = (certain - mean(certain)) / sd(lottery),
         
         # if matches the fixed dollar option, 0, else 1 (for risky)
         chose_risky = ifelse(str_detect(response, "Keep \\$[0-9]+") | 
                                str_detect(response, "Forfeit \\$[0-9]+"), 0, 1)) |>
  left_join(IUS |>
              pivot_longer(!id, names_to = "item", values_to = "response") |>
              summarize(.by = id,
                        IUS = mean(response)),
            by = "id") |>
  left_join(BIS |>
              pivot_longer(!id, names_to = "item", values_to = "response") |>
              summarize(.by = id,
                        BIS = mean(response)),
            by = "id") |>
  left_join(BAS |>
              pivot_longer(!id, names_to = "item", values_to = "response") |>
              summarize(.by = id,
                        BAS = mean(response)),
            by = "id") |>
  left_join(STAI |>
              pivot_longer(!id, names_to = "item", values_to = "response") |>
              summarize(.by = id,
                        STAI_T = mean(response)),
            by = "id") |>
  mutate(lottery_certain_diff = ev_lottery - ev_certain,
         across(c(IUS, BIS, BAS, STAI_T), ~ (.x - mean(.x)) / sd(.x),
                .names = "{.col}_z"))
```


## Modeling

```{r}
#| eval: false
m0 <- lme4::glmer(chose_risky ~ 1 + (1 | item) + (1 | id),
            dat, family = "binomial")
saveRDS(m0, here::here("Models", "risky-choice_0.rds"))

m1.0 <- lme4::glmer(chose_risky ~ IUS_z + BIS_z + BAS_z + STAI_T_z + lottery_certain_diff + frame + 
                      (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))
saveRDS(m1.0, here::here("Models", "risky-choice_1.0.rds"))

m1.1 <- lme4::glmer(chose_risky ~ IUS_z + BIS_z + BAS_z + STAI_T_z + lottery_certain_diff + frame + 
                      IUS_z:frame + BIS_z:frame + BAS_z:frame + STAI_T_z:frame + lottery_certain_diff:frame +
                      IUS_z:lottery_certain_diff + BIS_z:lottery_certain_diff + BAS_z:lottery_certain_diff + STAI_T_z:lottery_certain_diff +
                    (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))
saveRDS(m1.1, here::here("Models", "risky-choice_1.1.rds"))
```

::: {style="font-size:70%"}
``` {r}
#| layout-ncol: 3
m0 <- readRDS(here::here("Models", "risky-choice_0.rds"))
m1.0 <- readRDS(here::here("Models", "risky-choice_1.0.rds"))
m1.1 <- readRDS(here::here("Models", "risky-choice_1.1.rds"))

sjPlot::tab_model(m0,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR",
                 title = "Unconditional")
sjPlot::tab_model(m1.0,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR",
                 title = "No Interactions")
sjPlot::tab_model(m1.1,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR",
                 title = "Some Interactions")
```
:::

##### Interaction Effects

```{r}
#| layout-ncol: 3
#| fig-width: 3
#| fig-height: 3
#| message: false
effect_labels <- list("STAI_T_z[all]" = "STAI_T",
                      "BIS_z[all]" = "BIS_z",
                      "BAS_z[all]" = "BAS_z",
                      "frame" = "Frame",
                      "IUS_z[all]" = "IUS",
                      "lottery_certain_diff[all]" = "Lottery - Certain",
                      "lottery_certain_diff" = "Lottery - Certain")

sig_interactions <- list(c("STAI_T_z[all]", "frame"),
                         c("lottery_certain_diff[all]", "frame"),
                         c("IUS_z[all]", "lottery_certain_diff"),
                         c("BIS_z[all]", "lottery_certain_diff"),
                         c("BAS_z[all]", "lottery_certain_diff"))

sig_interactions |> walk(~ (ggeffects::ggpredict(m1.1, .x) |>
                              plot() +
                              labs(title = NULL, y = "P(Chose Risky)",
                                   x = effect_labels[.x[1]],
                                   fill = effect_labels[.x[2]],
                                   color = effect_labels[.x[2]]) +
                              coord_cartesian(ylim = c(0, 1)) +
                              theme_classic(base_size = 12) +
                              theme(aspect.ratio = 1,
                                    legend.position = "bottom")) |>
                           print())
```

##### Main Effects

```{r}
#| layout-ncol: 4
#| fig-width: 3
#| fig-height: 3
#| column: page
sig_mains <- list("BIS_z[all]", "BAS_z[all]", "lottery_certain_diff[all]", "frame")

sig_mains |> walk(~ (ggeffects::ggpredict(m1.1, .x) |>
                              plot() +
                              labs(title = NULL, y = "P(Chose Risky)",
                                   x = effect_labels[.x]) +
                              coord_cartesian(ylim = c(0, 1)) +
                              theme_classic(base_size = 12) +
                              theme(aspect.ratio = 1)) |>
                           print())
```

### Anxiety Scales Separate Models

```{r}
#| eval: false
m1.2.1 <- lme4::glmer(chose_risky ~ IUS_z * lottery_certain_diff * frame +
                    (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))
m1.2.2 <- lme4::glmer(chose_risky ~ STAI_T_z * lottery_certain_diff * frame +
                    (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))
m1.2.3 <- lme4::glmer(chose_risky ~ BIS_z * lottery_certain_diff * frame +
                    (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))
m1.2.4 <- lme4::glmer(chose_risky ~ BAS_z * lottery_certain_diff * frame +
                    (1 | item) + (1 | id), 
                  dat, family = "binomial",
                  lme4::glmerControl(optimizer = "bobyqa"))

saveRDS(m1.2.1, here::here("Models", "risky-choice_1.2.1.rds"))
saveRDS(m1.2.2, here::here("Models", "risky-choice_1.2.2.rds"))
saveRDS(m1.2.3, here::here("Models", "risky-choice_1.2.3.rds"))
saveRDS(m1.2.4, here::here("Models", "risky-choice_1.2.4.rds"))
```


::: {style="font-size:70%"}
```{r}
#| layout-ncol: 4
m1.2.1 <- readRDS(here::here("Models", "risky-choice_1.2.1.rds"))
m1.2.2 <- readRDS(here::here("Models", "risky-choice_1.2.2.rds"))
m1.2.3 <- readRDS(here::here("Models", "risky-choice_1.2.3.rds"))
m1.2.4 <- readRDS(here::here("Models", "risky-choice_1.2.4.rds"))

sjPlot::tab_model(m1.2.1,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR")

sjPlot::tab_model(m1.2.2,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR")

sjPlot::tab_model(m1.2.3,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR")

sjPlot::tab_model(m1.2.4,
                 show.ci = F,
                 show.se = T,
                 string.se = "SE",
                 show.r2 = F,
                 dv.labels = "",
                 string.est = "OR")
```
:::


```{r}
#| layout-ncol: 4
#| layout-nrow: 2
#| column: page
#| fig-width: 5
#| fig-height: 3
list(scale = list("IUS_z", "STAI_T_z", "BIS_z", "BAS_z"),
     model = list(m1.2.1, m1.2.2, m1.2.3, m1.2.4)) |>
  pwalk(\(scale, model) (ggeffects::ggpredict(model, terms = c("lottery_certain_diff[all]", scale, "frame")) |>
                              plot() +
                              labs(title = scale, y = "P(Chose Risky)") +
                              coord_cartesian(ylim = c(0, 1)) +
                              theme_classic(base_size = 12) +
                              theme(aspect.ratio = 1,
                                    legend.position = "bottom")) |>
         print())

list(scale = list("IUS_z", "STAI_T_z ", "BIS_z", "BAS_z"),
     model = list(m1.2.1, m1.2.2, m1.2.3, m1.2.4)) |>
  pwalk(\(scale, model) (ggeffects::ggpredict(model, terms = c(paste(scale, "[all]"), "frame")) |>
                              plot() +
                              labs(title = scale, y = "P(Chose Risky)") +
                              coord_cartesian(ylim = c(0, 1)) +
                              theme_classic(base_size = 12) +
                              theme(aspect.ratio = 1,
                                    legend.position = "bottom")) |>
         print())
```




